---
output: github_document
bibliography: bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# burger-emt-pdl1-models-figures

<!-- badges: start -->
<!-- badges: end -->

Models and code to generate figures for “Bidirectional Crosstalk Between Epithelial-Mesenchymal Plasticity and IFNγ-Induced PD-L1 Expression Promotes Tumor Progression.” bioRxiv. https://doi.org/10.1101/2022.02.03.478950 [@Burger2022_biorxiv].

## Usage

### Models

COPASI models can be found in `data/copasi_models`.

### Figures

Figures are generated reproducibly in R using [`renv`](https://rstudio.github.io/renv/index.html) and [`targets`](https://docs.ropensci.org/targets/):

1. Download/clone this repository
2. Open the project file (`.Rproj`) in RStudio
3. Run 
    ```{r, eval=F}
renv::restore()
```
    to install R package dependencies.

4. Open `targets.Rmd` and choose *Run* > *Run All*. Figures will appear in the `output` folder.

---