---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RefManageR)
bib <- ReadBib("bibliography.bib", check = FALSE)
```

# burger-emt-pdl1-models-figures

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6000437.svg)](https://doi.org/10.5281/zenodo.6000437)
<!-- badges: end -->

Models and code used to generate figures for

> `r NoCite(bib, c("Burger2022_rsos")); 
res <- capture.output(PrintBibliography(bib, .opts = list(style = "text", bib.style = "authoryear", no.print.fields = c("doi", "eprint", "issn", "isbn")))); gsub("(\\b[12][0-9]{3})NA\\b", "\\1", res)`


## Usage

### Models

COPASI models can be found in `data/copasi_models`.

### Figures

Figures are generated reproducibly in R using [`renv`](https://rstudio.github.io/renv/index.html) and [`targets`](https://docs.ropensci.org/targets/):

1. Download/clone this repository
2. Open the project file (`.Rproj`) in RStudio
3. Run 
    ```{r, eval=F}
    renv::restore()
    ```
    to install R package dependencies.

4. Open `targets.Rmd` and choose *Run* > *Run All*. Figures will appear in the `output` folder.
