---
title: "Figures for Bidirectional crosstalk between epithelial-mesenchymal plasticity and IFNγ-induced PD-L1 expression promotes tumor progression"
author: "Gerhard Burger & Carlijn Lems"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_chunk$set(tar_interactive = F)
options(scipen=999)
```

# Packages

```{r, eval = FALSE}
# DON'T USE use renv::restore
# also DESCRIPTION now handles dependencies for renv
install.packages("remotes")
remotes::install_github("ropensci/targets")
remotes::install_github("clauswilke/colorblindr")
remotes::install_cran(c("drc", "colorspace", "grid", "cowplot", "magick", "targets", "tarchetypes", "scales", "glue", "colorblindr", "fs", "rootSolve", 'tidyverse','interp'))
                      
source("R/functions.R")
source("R/common_plotstuff.R")
```

# Setup

First, load `targets` to activate the specialized `knitr` engine for Target Markdown.

```{r}
library(targets)

# my custom target utils
tar_mload <- function(names) {
  tar_make({{names}}) 
  tar_load({{names}}, envir = parent.frame(n = 2))
}
tar_mread <- function(name) {
  tar_mload({{name}})
  name
}
```

```{r}
tar_unscript()
```

# Globals

We first define some global options/functions common to all targets. The function below plots a histogram of ozone concentrations, and our histogram target will need it.

```{targets 00globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
options(scipen=999)
source("R/functions.R")
source("R/common_plotstuff.R")
tar_option_set(packages = c("drc", "cowplot", "magick", "grid",  "colorspace", "tarchetypes", "scales", "glue", "magrittr", "colorblindr", "fs", "rootSolve", "tidyverse", "interp"))
library(tarchetypes)

text_size <- 15
label_size <- round(text_size *1.5)
```

# Targets

## Figures

### 1. Model construction

```{targets figure1_pre}
list(
  # biorender svg is imported pdf in inkscape with poppler
  tar_target(jakstat_file, "data/biorender/biorender_jakstatpdl.svg", format = "file"),
  tar_target(model_file, "data/inkscape/TCSandPDL1.svg", format = "file"),
  tar_target(jakstat_bior, ggdraw() + draw_image(image_read(jakstat_file, density = 300))),
  tar_target(model_plot, ggdraw() + draw_image(image_read(model_file, density = 300)))
)
```

#### Time course data

```{targets figure1_timecourse}
list(
  tar_target(ifnstat_file, 'data/time_courses_sub/jakstat_time_course_ifn1.txt', format = "file"),
  tar_target(ifnpdl_file, 'data/time_courses_sub/ifn_jak_stat_pdl_timecourse_ifn1.txt', format = "file"),
  tar_target(ifnstat_timecourse, read_copasi(ifnstat_file)),
  tar_target(ifnpdl_timecourse, read_copasi(ifnpdl_file)),
  tar_target(stcs_timecourse, read_copasi(emt_no_int_file)),
  tar_target(ifnstat_tc_plot, 
             ggplot(ifnstat_timecourse) + 
               geom_line(aes(x = `# Time` / 60, y = `[x10]`), size = 1.1) +
               labs(x = "Time (minutes)", y = "STAT1p_2\n(nM)") + 
               theme_cowplot()),
  tar_target(ifnpdl_tc_plot, 
             ggplot(ifnpdl_timecourse) + 
               geom_line(aes(x = `# Time`, y = `[PM]`), size = 1.1) +
               scale_x_continuous(breaks = seq(0,48, by = 12), limits = c(0,48)) +
               labs(x = "Time (hours)", y = "PD-L1 membr.\n(molecules)") + 
               theme_cowplot()),
  tar_target(stcs_tc_plot, 
             ggplot(stcs_timecourse) + 
               geom_line(aes(x = `# Time` / 24, y = `[mZ]`), size = 1.1) +
               labs(x = "Time (days)", y = "ZEB1 mRNA\n(molecules)") + 
               theme_cowplot()),
  tar_target(tc_aligned, cowplot::align_plots(ifnstat_tc_plot, ifnpdl_tc_plot, stcs_tc_plot, align = "v", axis = 'l')),
  tar_target(tc_combined, plot_grid(tc_aligned[[1]], tc_aligned[[2]], tc_aligned[[3]], nrow = 1))
)
```

#### TCS model

```{targets tcs_bif}
list(
  tar_target(labels_p1, tribble(
    ~text, ~x, ~y,
    "E", 187000, 110,
    "E/M", 205000, 420,
    "M", 205000, 920
  ) %>% mutate(x = x / 1e5)),
  tar_target(stcs_data_wbp, 
             core %>%
               nest(data = everything()) %>%
               rowwise() %>%
               mutate(
                 bp = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]")),
                 data = list(data %>% mutate(interval = findInterval(`[mZ]`, bp)))
               ) %>%
               unnest(cols = data) %>%
               mutate(stable = (interval %% 2) == 0)),
  tar_target(stcs_plot, 
             stcs_data_wbp %>%
               ggplot(aes(`# Values[S].InitialValue`/1e5, `[mZ]`, group = interval, 
                          linetype = stable)) + 
               geom_line(size = 1.2) +
               scale_linetype_discrete(guide = "none", limits = c(T,F)) +
               labs(x = bquote('SNAIL1 ('*10^5~'molecules)'), y = 'ZEB1 mRNA (molecules)') +
               geom_text(data = labels_p1, aes(x,y, label = text), 
                         fontface = "bold", inherit.aes = F, show.legend = FALSE) + 
               coord_cartesian(xlim = c(1.8, 2.3), ylim = c(0, 1000)) +
               scale_x_continuous(expand = c(0,0)) +
               scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
               expand_limits(y = 0) +
               theme_cowplot() + 
               theme(legend.position = "top"))
)
```

```{targets tcs_tc}
tcs_tc_path <- "data/time_courses_tcs"
list(
  tar_target(E_tc_file,  path(tcs_tc_path, "core_emt_E200000_to_S215000.txt"), format = "file"),
  tar_target(EM_tc_file, path(tcs_tc_path, "core_emt_EM200000_to_S215000.txt"), format = "file"),
  tar_target(M_tc_file,  path(tcs_tc_path, "core_emt_M200000_to_S215000.txt"), format = "file"),
  tar_target(E_ss_file,  path(tcs_tc_path, "core_emt_E200000_to_S215000_SS.txt"), format = "file"),
  tar_target(EM_ss_file, path(tcs_tc_path, "core_emt_EM200000_to_S215000_SS.txt"), format = "file"),
  tar_target(M_ss_file,  path(tcs_tc_path, "core_emt_M200000_to_S215000_SS.txt"), format = "file"),
  
  tar_target(tc_data, tibble(emt = c("E", "E/M", "M"), 
                             file = c(E_tc_file, EM_tc_file, M_tc_file)) %>%
               mutate(emt = factor(emt, levels = c("M", "E/M", "E"))) %>%
               group_by(emt) %>%
               rowwise() %>%
               summarise(data = list(read_copasi(file)), .groups = "drop") %>% 
               unnest(data)),
  tar_target(ss_data, tibble(emt = c("E", "E/M", "M"), 
                             file = c(E_ss_file, EM_ss_file, M_ss_file)) %>%
               mutate(emt = factor(emt, levels = c("M", "E/M", "E"))) %>%
               group_by(emt) %>%
               rowwise() %>%
               summarise(data = list(read_copasi(file) %>% 
                                       arrange(`# Values[S].InitialValue`)), 
                         .groups = "drop") %>% 
               unnest(data)),
  tar_target(ss_data_min, ss_data %>% group_by(emt) %>%
               filter(`# Values[S].InitialValue` == min(`# Values[S].InitialValue`))),
  tar_target(ss_data_max, ss_data %>% group_by(emt) %>%
               filter(`# Values[S].InitialValue` == max(`# Values[S].InitialValue`))),
  tar_target(tc_plot, 
             ggplot(tc_data, aes(`# Time` / 24, `[mZ]`, color = emt)) + 
               geom_line(size = 1.2, show.legend = F) +
               labs(x = "Time (days)", y = "ZEB mRNA (molecules)") +
               coord_cartesian(ylim = c(0, 1000)) +
               scale_x_continuous(expand = c(0,0)) +
               scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
               expand_limits(y = 0) +
               theme_cowplot()),
  tar_target(tc_ss_plot, 
             stcs_plot +   
               geom_point(data = ss_data_min, aes(`# Values[S].InitialValue`, `[mZ]`, color = emt),
                          size = 3, inherit.aes = F, show.legend = F) +
               geom_line(data = ss_data, aes(`# Values[S].InitialValue`, `[mZ]`, color = emt), 
                         size = 1.2, inherit.aes = F, show.legend = F, 
                         arrow = arrow(length = unit(0.1, "inches"), type = "closed"))),
  tar_target(tc_tot_aligned, cowplot::align_plots(stcs_plot, tc_ss_plot, 
                                                  align = "v", axis = 'l'))
)
```
```{targets tc_ss_alt_plot, tar_simple = T}
ss_data_minmax <- left_join(
  ss_data_max %>% mutate(S = 215000, mZ_high = `[mZ]`) %>% select(emt, starts_with("mZ")),
  ss_data_min %>% mutate(S = 200000, mZ_low = `[mZ]`) %>% select(emt, starts_with("mZ")),
  by = "emt"
) %>% 
  mutate(emt = factor(emt, levels = c("M", "E/M", "E")))

stcs_plot +   
  geom_point(data = ss_data_min, aes(2, `[mZ]`, color = emt),
             size = 2.5, inherit.aes = F, show.legend = F) + 
  geom_point(data = ss_data_min, aes(2.15, `[mZ]`, color = emt),
             size = 2.5, inherit.aes = F, show.legend = F) +
  geom_segment(data = ss_data_min, aes(x = 2, y = `[mZ]`, color = emt,
                                       xend = 2.15, yend = `[mZ]`),
             size = 1.2, inherit.aes = F, show.legend = F, linetype = "dashed") +
  geom_segment(data = ss_data_minmax, aes(x = 2.15, y = mZ_low, color = emt,
                                       xend = 2.15, yend = mZ_high),
             size = 1.2, arrow = arrow(length = unit(0.1, "inches"), type = "closed"), 
             inherit.aes = F, show.legend = F) + 
  geom_text(data = labels_p1, aes(x,y, label = text), 
            fontface = "bold", inherit.aes = F, show.legend = FALSE)  
```


#### Combined

```{targets figure1}
list(
  tar_target(figure1, {
    plot_grid(
      plot_grid(jakstat_bior, model_plot,
                nrow = 1, rel_widths = c(1.1,2),
                labels = c("A", "B"), label_size = label_size),
      NULL, NULL,
      tc_combined,
      plot_grid(NULL,
                tc_ss_alt_plot,
                NULL,
                tc_plot,
                NULL,
                nrow = 1, align = "h", rel_widths = c(0.05, 1,0.05,1,1),
                labels = c("D","", "", "", ""), label_size = label_size),
      ncol = 1, rel_heights = c(2.5, 0.05, 0.1, 1,1.7),
      labels = c("", "", "C", ""), label_size = label_size
    )
  }),
  tar_target(figure1_file, {
    path <- "output/figure1.pdf"
    ggsave(path, plot = figure1, width = 12, height = 9, device = cairo_pdf)
  })
)

```

### 2. IFNy -> EMT, EMT -> PD-L1

```{targets figure2-source-files}
bifs_path <- "data/bifs"
list(
  tar_target(twod_copasi_file, path(bifs_path, "full_model_2d_bif.txt.xz"), format = "file"),
  tar_target(core_emt_file, path(bifs_path, "core_emt.txt"), format = "file"),
  tar_target(ifn_bif_file, path(bifs_path, "full_model_ifn_bif.txt.xz"), format = "file"),
  tar_target(twod_copasi, read_copasi(twod_copasi_file)),
  tar_target(core, read_copasi(core_emt_file)),
  tar_target(g_I_cont, read_copasi(ifn_bif_file))
)
```
#### Snail bifurcation

```{targets bif_globals}
list(
  tar_target(xmin, 170000 / 1e5),
  tar_target(xmax, (230000+1) / 1e5),
  tar_target(exp_order, c("no_int", "ifn0", "ifn1")),
  tar_target(exp_labels,  c("TCS", 
                expression(paste('IFN', gamma, ' = 0 nM')),
                expression(paste('IFN', gamma, ' = 0.1 nM'))
                ) %>% setNames(exp_order)) 
)
```

```{targets bifs_snail_data, tar_simple = T}
data <- tribble(
  ~experiment, ~data,
  "no_int", core,
  "ifn0", twod_copasi %>% filter(`Values[I (global)].InitialValue` == 0),
  "ifn1", twod_copasi %>% filter(`Values[I (global)].InitialValue` == 0.1),
) %>% 
  rowwise() %>%
  mutate(
    bp = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]")),
    data = list(data %>% mutate(interval = findInterval(`[mZ]`, bp)))
  ) %>%
  unnest(cols = data) %>%
  mutate(stable = (interval %% 2) == 0)

data %>% mutate(experiment = factor(experiment, levels = exp_order))
```


```{targets bif_snail_zeb, tar_simple = T}
labels_p1 <- tribble(
  ~text, ~x, ~y,
  "E", 177000, 90,
  "E/M", 197000, 410,
  "M", 205000, 920
) %>% mutate(x = x / 1e5)

bifs_snail_data %>%
  ggplot(aes(`# Values[S].InitialValue` / 1e5, `[mZ]`, group = interaction(experiment, interval), linetype = stable, color = experiment)) + 
  geom_line(size = 1.2) + 
  geom_vline(xintercept = 190000 / 1e5, linetype = "dashed") +
  scale_linetype_discrete(guide = "none", limits = c(T,F)) +
  scale_color_OkabeIto(order = c(8,2,1), labels = exp_labels, name = NULL) + 
  labs(x = bquote('SNAIL1 ('*10^5~'molecules)'), y = 'ZEB1 mRNA (molecules)') +
    geom_text(data = labels_p1, aes(x,y, label = text), fontface = "bold", inherit.aes = F, show.legend = FALSE) + 
  coord_cartesian(xlim = c(xmin,xmax)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) + 
  expand_limits(y = 0) +
  theme(legend.position = "top")
```

```{targets bif_snail_pdl, tar_simple = T}
labels_p2 <- tribble(
  ~experiment, ~text, ~x, ~y,
  "ifn0", "E", 180000, 24000,
  "ifn0", "E/M", 211000, 16000,
  "ifn0", "M", 225000, 67000,
  "ifn1", "E", 175000, 65000,
  "ifn1", "E/M", 200000, 115000,
  "ifn1", "M", 210000, 310000
) %>% mutate(x = x / 1e5)

bifs_snail_data %>%
  filter(experiment != "no_int") %>%
  ggplot(aes(`# Values[S].InitialValue`/ 1e5, `[PM]`, group = interaction(experiment, interval), linetype = stable, color = experiment)) + 
  geom_line(size = 1.2) + 
  geom_vline(xintercept = 190000 / 1e5, linetype = "dashed") + 
  scale_linetype_discrete(guide = "none", limits = c(T,F)) +
  scale_color_OkabeIto(order = c(2,1), labels = exp_labels, name = NULL) + 
  labs(x = bquote('SNAIL1 ('*10^5~'molecules)'), y = 'PD-L1 membrane (molecules)') +
  geom_text(data = labels_p2, aes(x,y, label = text, color = experiment), fontface = "bold", inherit.aes = F, show.legend = FALSE) + 
  coord_cartesian(xlim = c(xmin,xmax), ylim = c(0, 3.2e5)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) + 
  theme(legend.position = "top")
```


#### IFN bifurcation

```{targets bif_I_P, tar_simple = T}
mZ_breaks_g_I_cont <- determine_BP(g_I_cont, "# Values[I (global)].InitialValue", "[mZ]")

g_I_cont %>% mutate(
  interval = findInterval(`[mZ]`, mZ_breaks_g_I_cont),
  stable = (interval %% 2) == 0
) -> g_I_cont_pd

labels_bifIP <- tribble(
  ~text, ~x, ~y,
  "E", 0.075, 0.45e5,
  "E/M", 0.075, 1.05e5,
  "M", 0.075, 2.65e5
)

g_I_cont_pd %>%
  ggplot(aes(`# Values[I (global)].InitialValue`, `[PM]`, group = interval, linetype = stable)) + 
  geom_line(size = 1.2) + 
  scale_linetype_discrete(guide = "none", limits = c(T,F)) + 
  coord_cartesian(xlim = c(0, 0.1), ylim = c(0, 3.2e5)) +
  geom_text(data = labels_bifIP, aes(x,y, label = text), fontface = "bold", 
            inherit.aes = F, show.legend = FALSE) + 
  xlab(expression(paste('IFN', gamma, ' (nM)'))) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) + 
  ylab('PD-L1 membrane (molecules)')
```
#### 2D bif

```{targets fixed_bps_twod_copasi, tar_simple = T}
twod_copasi %>% 
  drop_na() %>%
  nest_by(`Values[I (global)].InitialValue`) -> nested

# Find bifurcations x and y points
nested <- nested %>%
  mutate(
    bpx = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]", res = "x")),
    bpy = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]", res = "y")),
  )

# sometimes too many bps are detected
wonky <- nested %>% filter(length(bpx) != 4)

## Inspect the wonky bps
# wonky %>% mutate(
#   data = list(data %>% mutate(interval = findInterval(`[mZ]`, bpy)))
# ) %>% 
#   select(-bpx, -bpy) %>%
#   unnest(data) %>% 
#   ggplot(aes(`# Values[S].InitialValue`,`[mZ]`, color = as.factor(interval))) + 
#            geom_point() +
#            facet_wrap(vars(`Values[I (global)].InitialValue`))

correct_bps <- tibble(
  `Values[I (global)].InitialValue` = wonky$`Values[I (global)].InitialValue`,
  bps = list(
    c(1:4),
    c(1,2,5,6),
    c(1:4),
    c(1,2,3,6),
    c(1,2,3,6),
    c(1,2,5,6),
    c(1,2,3,6),
    c(1,2,3,6),
    c(1,2,3,6),
    c(1,2,5,6)
  )
)
  
fixed <- wonky %>% 
  left_join(correct_bps) %>%
  rowwise() %>%
  mutate(
    bpx = list(bpx[bps]),
    bpy = list(bpy[bps])
  ) %>%
  select(-bps)

## check fixed bps
# fixed %>% mutate(
#   data = list(data %>% mutate(interval = findInterval(`[mZ]`, bpy)))
# ) %>% 
#   select(-bpx, -bpy) %>%
#   unnest(data) %>% 
#   ggplot(aes(`# Values[S].InitialValue`,`[mZ]`, color = as.factor(interval))) + 
#   geom_point() +
#   facet_wrap(vars(`Values[I (global)].InitialValue`))

nested %>%
  rows_update(fixed)
```

```{targets twod_copasi_bps_res}
list(
  tar_target(twod_bps_raw, {
    fixed_bps_twod_copasi %>% 
      summarize(bp_to_tibble(bpx), .groups = "drop") %>%
      rename(IFN = `Values[I (global)].InitialValue`)
  }),
  tar_target(twod_bps_smooth, {
    BP1_sp <- spline(twod_bps_raw$IFN, twod_bps_raw$BP1, xmax = 0.1, n = 10*length(twod_bps_raw$IFN))
    BP2_sp <- spline(twod_bps_raw$IFN, twod_bps_raw$BP2, xmax = 0.1, n = 10*length(twod_bps_raw$IFN))
    BP3_sp <- spline(twod_bps_raw$IFN, twod_bps_raw$BP3, xmax = 0.1, n = 10*length(twod_bps_raw$IFN))
    BP4_sp <- spline(twod_bps_raw$IFN, twod_bps_raw$BP4, xmax = 0.1, n = 10*length(twod_bps_raw$IFN))
BP_smooth <- tibble(IFN = BP1_sp$x, BP1 = BP1_sp$y, BP2 = BP2_sp$y, BP3 = BP3_sp$y, BP4 = BP4_sp$y)
  })
)
```



```{targets init_model_pd, tar_simple = T}
ggplot(twod_bps_smooth) +
  geom_ribbon(aes(y = IFN, xmin = xmin, xmax = xmax, fill = '{E}')) +
  geom_ribbon(aes(y = IFN, xmin = BP4 / 1e5, xmax = xmax, fill = '{E, M}')) +
  geom_ribbon(aes(y = IFN, xmin = BP2 / 1e5, xmax = xmax, fill = '{E, E/M, M}')) +
  geom_ribbon(aes(y = IFN, xmin = BP1 / 1e5, xmax = xmax, fill = '{E/M, M}')) +
  geom_ribbon(aes(y = IFN, xmin = BP3 / 1e5, xmax = xmax, fill = '{M}')) +
  geom_vline(xintercept = 190000 / 1e5, linetype = "dashed") +
  coord_cartesian(ylim = c(0,0.1), expand = F) +
  labs(x = bquote('SNAIL1 ('*10^5~'molecules)'), y = expression(paste('IFN', gamma, ' (nM)')), fill = NULL) +
  #scale_fill_manual(name = "", values = c('{E, E/M, M}' = custom.col[15], '{E, M}' = custom.col[4], '{E/M, M}' = custom.col[13], '{E}' = custom.col[6], '{M}' = custom.col[24]), limits = c('{E}', '{E, M}', '{E, E/M, M}', '{E/M, M}', '{M}')) +
  scale_fill_OkabeIto(order = 3:7, limits = c('{E}','{E, M}', '{E, E/M, M}', '{E/M, M}', '{M}')) + 
  scale_y_continuous(breaks = seq(0, 0.1, by =0.02))  
```

#### Combined

```{targets figure2}
list(
  tar_target(figure2, {
    plot_grid(bif_snail_pdl + theme_cowplot(text_size)+ theme(legend.position="top", legend.justification = "right"), 
              bif_I_P + theme_cowplot(text_size)+ theme(legend.position="top"), 
              bif_snail_zeb + theme_cowplot(text_size)+ theme(legend.position="top", legend.justification = "right"), 
              init_model_pd + theme_cowplot(text_size) + theme(legend.position="top", legend.justification = "center"), 
              labels = "AUTO", label_size = label_size, ncol = 2, align = "vh", axis = "tlbr",
              scale = 0.99) #< allow spacing for labels
  }),
  tar_target(figure2_file, {
    path <- "output/figure2.pdf"
    ggsave(path, plot = figure2, width = 12, height = 10, device = cairo_pdf)
  })
)
```


### 3. Temporal dynamics

```{targets temp_dyn_files}
emt_td_path <- "data/emt_td"
list(
  tar_target(emt_no_int_file, path(emt_td_path, "EMT no interaction.txt"), format = "file"),
  tar_target(emt_ifn0_file, path(emt_td_path, "EMT no IFN fixed.txt"), format = "file"),
  tar_target(emt_ifn1_file, path(emt_td_path, "EMT IFN fixed.txt"), format = "file"),
  tar_target(met_no_int_file, path(emt_td_path, "MET no interaction.txt"), format = "file"),
  tar_target(met_ifn0_file, path(emt_td_path, "MET no IFN fixed.txt"), format = "file"),
  tar_target(met_ifn1_file, path(emt_td_path, "MET IFN fixed.txt"), format = "file")
)
```
```{targets time_dynamics_long, tar_simple = T}
time_dynamics <- tribble(
  ~type, ~experiment, ~file,
  "EMT", "no_int", emt_no_int_file,
  "EMT", "ifn0", emt_ifn0_file,
  "EMT", "ifn1", emt_ifn1_file,
  "MET", "no_int", met_no_int_file,
  "MET", "ifn0", met_ifn0_file,
  "MET", "ifn1", met_ifn1_file,

) %>% 
  rowwise() %>%
  mutate(data = list(read_copasi(file))) %>%
  select(-file) %>% unnest(data)

time_dynamics <- time_dynamics %>%
  mutate(
    across(all_of(c("[mP]", "[PM]","(R5).Flux")), ~ ifelse(experiment == "no_int", NA, .x)),
    time = `# Time`/24
  ) %>% select(-`# Time`)

time_dynamics %>% 
  pivot_longer(cols = c(where(is.numeric), -time))
```
```{targets td_gvars}
list(
  tar_target(td_vars, c("[mu2]", "[mZ]", "[Z]", "[mP]", "[PE]", "[PG]", "[PM]", 
                        "[mF]", "[F]")),
  tar_target(td_labels, c(
    "miR-200",
    "ZEB1 mRNA ",
    "ZEB1",
    "PD-L1 mRNA",
    "PD-L1 ER",
    "PD-L1 Golgi",
    "PD-L1 membrane",
    "IRF1 mRNA",
    "IRF1"
  ) %>% setNames(td_vars)),
  tar_target(td_vars_to_show, c("[mu2]", "[mZ]", "[Z]", "[mP]", "[PM]"))
)
```


```{targets figure3, tar_simple = T}
time_dynamics_long %>% 
  filter(name %in% td_vars_to_show) %>%
  mutate(experiment = factor(experiment, levels = exp_order)) %>%
  ggplot(aes(time,value, color = experiment)) +
  geom_line(aes(linetype = experiment), size = 1.2) +
  facet_grid(rows = vars(factor(name, levels = td_vars)), cols = vars(type), 
             scales = "free_y", switch = "y",
             labeller = labeller(.rows = as_labeller(td_labels))) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  expand_limits(y = 0) + 
  scale_color_OkabeIto(order = c(8,2,1), labels = exp_labels, name = NULL) +
  scale_linetype_manual(values = c("dashed", "solid", "solid"), labels = exp_labels, name = NULL) + 
  labs(x = "Time (days)", y = "Regulator (molecules)") + 
  theme_cowplot(text_size) + 
  panel_border() +
  #background_grid() + 
  theme(strip.placement = 'outside', 
        strip.background = element_blank(),
        strip.text.x = element_text(size = 1.2*text_size, face = "bold"),
        panel.spacing.x = unit(1, "lines"),
        legend.position = "top",
        legend.justification = "right")
```

#### Combined

```{targets fig3}
  tar_target(figure3_file, {
    path <- "output/figure3.pdf"
    ggsave(path, plot = figure3, width = 12, height = 10, device = cairo_pdf)
  })
```


## SI

### S1. STAT-PDL model

```{targets statpdl_model}
list(
  tar_target(statpdl_model_file, "data/inkscape/statpdl1.svg", format = "file"),
  tar_target(statpdl_model_fig, 
             ggdraw() + draw_image(image_read(statpdl_model_file, density = 300)))
)
```

```{targets statpdl_tc_plot, tar_simple = T}
vars <- c( "[mP]", "[PE]", "[PG]", "[PM]", "[mF]", "[F]")
labels <- c(
  "PD-L1 mRNA",
  "PD-L1 ER",
  "PD-L1 Golgi",
  "PD-L1 membrane",
  "IRF1 mRNA",
  "IRF1 protein"
) %>% setNames(vars)
var_order <- c("[mF]", "[F]", "[mP]", "[PE]", "[PG]", "[PM]")

ifnpdl_timecourse %>%
  select(-`[STAT]`) %>%
  mutate(time = `# Time`) %>% 
  select(-`# Time`) %>%
  pivot_longer(cols = c(where(is.numeric), -time)) %>%
  ggplot(aes(time,value)) +
  geom_line(size = 1.2) +
  facet_wrap(vars(factor(name, levels = var_order)), ncol = 2,
             scales = "free_y",
             labeller = as_labeller(labels)) +
  scale_x_continuous(breaks = seq(0,48, by = 12), limits = c(0,48), expand = c(0,0)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  expand_limits(y = 0) + 
  labs(x = "Time (hours)", y = "(molecules)") + 
  theme_cowplot(text_size) + 
  panel_border() +
  #background_grid() + 
  theme(strip.background = element_blank())
```

#### Combined

```{targets si_statpdl_model, tar_simple = T}
plot_grid(statpdl_model_fig, statpdl_tc_plot, 
          nrow = 1, rel_widths = c(1, 2),
          labels = "AUTO", label_size = label_size)
```

```{targets si_statpdl_model_file, tar_simple = T}
path <- "output/si_statpdl_model.pdf"
ggsave(path, plot = si_statpdl_model, width = 8, height = 4, device = cairo_pdf)
```

### S2. PD-L1 for different EMT states

Smooth and interpolate data to be able to use geom_raster

```{targets interped_data, tar_simple = T}

# Add EMT status to data based on BP interval
fixed_bps_twod_copasi %>%
  mutate(
    data = list(data %>% mutate(interval = findInterval(`[mZ]`, bpy)))
  ) %>%
  select(-bpx,-bpy) %>%
  unnest(data) %>%
  filter((interval %% 2) == 0 ) %>%
  left_join(tibble(interval = c(0,2,4), emt = c("E", "E/M", "M")), by = "interval") %>%
  select(-interval) %>%
  rename(
    IFN = `Values[I (global)].InitialValue`,
    S = `# Values[S].InitialValue`
    ) -> sumtest

# take average for duplicates
sumtest %>% 
  select(IFN, S, emt, `[PM]`) %>%
  group_by(IFN, S, emt) %>% 
  summarise(across(where(is.numeric), mean), .groups = "drop") -> simple_sum

# reduce x resolution to solve missing values and make suitable for interp
xintervals <- with(simple_sum, seq(min(S), max(S), length = 101))
xmidinterval <- xintervals + diff(xintervals)[1]
simple_sum %>% 
  ungroup() %>% 
  mutate(s_bin = findInterval(S, xintervals)) %>% 
  group_by(IFN, s_bin, emt) %>% 
  summarize(`[PM]` = mean(`[PM]`), .groups = "drop") %>%
  left_join(tibble(s_bin = seq_along(xintervals), s_bin_mid = xmidinterval), by = "s_bin") %>%
  mutate(S = s_bin_mid) %>%
  select(-starts_with("s_bin")) -> simple_sum2

simple_sum2 %>% 
  nest_by(emt) %>%
  mutate(
    data = list({
        interp(data$S / 1e5, data$IFN, data$`[PM]`, 
               xo = xmidinterval / 1e5, yo = unique(data$IFN)) %>% 
        interp2xyz(data.frame = T) %>% 
        rename(S = x, IFN = y, `[PM]` = z) %>%
        as_tibble() %>% 
        drop_na()
    }))-> interped_data_full

bounds <- tibble(
  emt = c("E", "E/M", "M"), 
  bounds = list(
    twod_bps_raw %>% select(IFN, s_max = BP1) %>% add_column(s_min = 0),
    twod_bps_raw %>% select(IFN, s_min = BP2, s_max = BP3),
    twod_bps_raw %>% select(IFN, s_min = BP4) %>% add_column(s_max = Inf)
  )
)

# remove interpolation outside hull
interped_data_full %>% 
  left_join(bounds, by = "emt") %>% 
  mutate(
    data= list(data %>% 
                 left_join(bounds, by = "IFN") %>%
                 filter(S >= s_min / 1e5, S <= s_max / 1e5) %>%
                 select(-s_min, -s_max)
    )
    ) %>% 
  select(-bounds) %>%
  unnest(data)
```

```{targets si_pdl_emt_plot_data, tar_simple = T}
interped_data %>%
  group_by(IFN, S) %>%
  summarize(`[PM]` = mean(`[PM]`, na.rm =T), .groups = "drop") -> averaged

bind_rows(interped_data, averaged %>% add_column(emt = "Average")) %>%
  mutate(emt = factor(emt, levels = c("E", "E/M", "M", "Average")))
```
```{targets si_pdl_emt_plot, tar_simple=T}
si_pdl_emt_plot_data %>%
  ggplot(aes(S,IFN)) + 
  geom_raster(aes(fill = `[PM]` / 1e5)) + 
  facet_wrap(vars(emt), nrow = 1) +
  scale_fill_viridis_c(limits = function(x) {x[1] <- 0; x}) +
  labs(x = bquote('SNAIL1 ('*10^5~'molecules)'),
       y = expression(paste('IFN', gamma, ' (nM)')),
       fill = bquote('PD-L1 membrane ('*10^5~'molecules)')) + 
  geom_path(data = twod_bps_smooth %>% pivot_longer(cols = starts_with("BP")), 
            aes(x = value / 1e5, y = IFN, group = name), 
            inherit.aes = F, size = 1, color = "white") +
  coord_cartesian(expand = F) + 
  scale_y_continuous(breaks = seq(0, 0.1, by =0.02)) +  
  theme_cowplot(text_size) + 
  panel_border() +
  theme(strip.background = element_blank()) + 
  theme(legend.position="top", 
        legend.justification = "right") 
```



#### Combined

```{targets si_pdl_emt_output}
list(
  tar_target(si_pdl_emt_file, {
    path <- "output/si_pdl_emt.pdf"
    ggsave(path, plot = si_pdl_emt_plot, width = 12, height = 4, device = cairo_pdf)
  })
)
```



### S3. Simplified model

```{targets si_simplified}
list(
  tar_target(simp_model_file, "data/inkscape/TCSandPDL1_wo_jakstat.svg", format = "file"),
  tar_target(simple_model_2d_bif_file, "data/bifs/core_emt_plus_pdl_2d_bif.txt.xz", format = "file"),
  tar_target(simple_model_2d_bif, read_copasi(simple_model_2d_bif_file))
)
```

```{targets si_simp_2d_bif, tar_simple = T}
simple_model_2d_bif %>% 
  drop_na() %>%
  nest_by(`Values[F].InitialValue`) -> nested

nested <- nested %>%
  mutate(
    bps = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]", res = "x"))
  )

nested %>% 
  filter(length(bps) <= 4) %>% # sometimes too many bps are detected
  summarize(bp_to_tibble(bps), .groups = "drop") %>%
  rename(FF = `Values[F].InitialValue`) -> sum2

spline_xmax <- 3e5
BP1_sp <- spline(sum2$FF, sum2$BP1, xmax = spline_xmax, n = 10*length(sum2$FF))
BP2_sp <- spline(sum2$FF, sum2$BP2, xmax = spline_xmax, n = 10*length(sum2$FF))
BP3_sp <- spline(sum2$FF, sum2$BP3, xmax = spline_xmax, n = 10*length(sum2$FF))
BP4_sp <- spline(sum2$FF, sum2$BP4, xmax = spline_xmax, n = 10*length(sum2$FF))
BP_smooth <- tibble(FF = BP1_sp$x, BP1 = BP1_sp$y, BP2 = BP2_sp$y, BP3 = BP3_sp$y, BP4 = BP4_sp$y)

p2 <- ggplot(BP_smooth) +
  geom_ribbon(aes(y = FF, xmin = xmin, xmax = xmax, fill = '{E}')) +
  geom_ribbon(aes(y = FF, xmin = BP4 / 1e5, xmax = xmax, fill = '{E, M}')) +
  geom_ribbon(aes(y = FF, xmin = BP2 / 1e5, xmax = xmax, fill = '{E, E/M, M}')) +
  geom_ribbon(aes(y = FF, xmin = BP1 / 1e5, xmax = xmax, fill = '{E/M, M}')) +
  geom_ribbon(aes(y = FF, xmin = BP3 / 1e5, xmax = xmax, fill = '{M}')) +
  coord_cartesian(ylim = c(0,300000)) +
  labs(x = bquote('SNAIL1 ('*10^5~'molecules)'), y = "I (molecules)", fill = NULL) +
  #scale_fill_manual(name = "", values = c('{E, E/M, M}' = custom.col[15], '{E, M}' = custom.col[4], '{E/M, M}' = custom.col[13], '{E}' = custom.col[6], '{M}' = custom.col[24]), limits = c('{E}', '{E, M}', '{E, E/M, M}', '{E/M, M}', '{M}')) +
  scale_fill_OkabeIto(order = 3:7, limits = c('{E}','{E, M}', '{E, E/M, M}', '{E/M, M}', '{M}')) + 
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_cowplot(text_size)
```

#### Combined

```{targets si_simp_output}
list(
  tar_target(si_simp_output,
    plot_grid(ggdraw() + draw_image(image_read(simp_model_file, density = 300)), 
              si_simp_2d_bif + theme(legend.position="top", legend.justification = "center"),
          nrow = 1, rel_widths = c(1,1),
          labels = c("A", "B"), label_size = label_size, scale = 0.90)
  ),
  tar_target(si_simp_file, {
    path <- "output/si_simplified_model.pdf"
    ggsave(path, plot = si_simp_output, width = 12, height = 4, device = cairo_pdf)
  })
)
```

### S4. Sensitivity

```{targets sens_files}
list(
  tar_files_input(sens_files_sep,
                  fs::path("data/sensitivity/",
                           c("lmp2 0.16.txt.xz",
                             "lmp2 0.18.txt.xz",
                             "lmp2 0.22.txt.xz",
                             "lmp2 0.24.txt.xz",
                             "lmp3 0.8.txt.xz",
                             "lmp3 0.9.txt.xz",
                             "lmp3 1.1.txt.xz",
                             "lmp3 1.2.txt.xz",
                             "lmump2 0.04.txt.xz",
                             "lmump2 0.045.txt.xz",
                             "lmump2 0.055.txt.xz",
                             "lmump2 0.06.txt.xz",
                             "lmump3 0.8.txt.xz",
                             "lmump3 0.9.txt.xz",
                             "lmump3 1.1.txt.xz",
                             "lmump3 1.2.txt.xz",
                             "ltp1 0.8.txt.xz",
                             "ltp1 0.9.txt.xz",
                             "ltp1 1.1.txt.xz",
                             "ltp1 1.2.txt.xz",
                             "ltp2 0.24.txt.xz",
                             "ltp2 0.27.txt.xz",
                             "ltp2 0.33.txt.xz",
                             "ltp2 0.36.txt.xz",
                             "ltp3 0.04.txt.xz",
                             "ltp3 0.045.txt.xz",
                             "ltp3 0.055.txt.xz",
                             "ltp3 0.06.txt.xz",
                             "mu2_0 lmp 11000.txt.xz",
                             "mu2_0 lmp 12000.txt.xz",
                             "mu2_0 lmp 8000.txt.xz",
                             "mu2_0 lmp 9000.txt.xz",
                             "mu2_0 lmump 11000.txt.xz",
                             "mu2_0 lmump 12000.txt.xz",
                             "mu2_0 lmump 8000.txt.xz",
                             "mu2_0 lmump 9000.txt.xz",
                             "mu2_0 ltp 11000.txt.xz",
                             "mu2_0 ltp 12000.txt.xz",
                             "mu2_0 ltp 8000.txt.xz",
                             "mu2_0 ltp 9000.txt.xz")
                  )
  ),
  tar_target(sens_file_unchanged, "data/sensitivity/unchanged.txt.xz", format = "file")
)
```
```{targets sens_tibble}
list(
  tar_target(meta, {
    meta <- tibble(file = sens_files_sep) %>%

      # extract parameter and value from file name
      mutate(str_match(path_file(file), "(.*) ([\\d\\.]+[\\d])(\\.txt)?")[,c(2,3)] %>% 
               `colnames<-`(c("par", "val")) %>%
               as_tibble()) %>% 
      mutate(val = as.numeric(val)) %>% 
      # sort before adding change
      group_by(par) %>%
      arrange(val, .by_group = T) %>%
      # add change (every parameter has -20%, -10%, +10%, +20%)
      mutate(change = rep_along(val, c(-0.2, -0.1, 0.1, 0.2))) %>%
      # add unchanged.txt
      ungroup() %>%
      add_row(tibble_row(file = sens_file_unchanged, par = "base", change = 0))
  }),
  tar_target(all_nested, {
    meta %>%
      # read data
      rowwise() %>%
      mutate(data = list(read_copasi(file))) %>%
      select(-file) %>%
      # computate bifurcation
      mutate(bp = list(determine_BP(data, "# Values[S].InitialValue", "[mZ]"))) %>%
      # include bp info in data
      mutate(data = list(data %>% mutate(interval = findInterval(`[mZ]`, bp)))) %>%
      select(-bp)
  }),
  tar_target(par_meta, {
    par_meta <- tribble(
      ~par, ~col, ~row,
      "lmp2", "mRNA_PD-L1", "l2",
      "lmp3", "mRNA_PD-L1", "l3",
      "lmump2", "miR-200", "l2",
      "lmump3", "miR-200", "l3",
      "ltp1", "PD-L1_ER", "l1",
      "ltp2", "PD-L1_ER", "l2",
      "ltp3", "PD-L1_ER", "l3",
      "mu2_0 lmp", "mRNA_PD-L1", "mu0", 
      "mu2_0 lmump", "miR-200" ,"mu0",
      "mu2_0 ltp", "PD-L1_ER", "mu0"
    )
    
    # l1, l2, and l3, are really l0, l1, and l2
    row_names <- c(paste0("l", 1:3), "mu0")
    row_labels <- c(
      expression(gamma[mu*0]/gamma["m"*0]/"l"[0]),
      expression(gamma[mu*1]/gamma["m"*1]/"l"[1]),
      expression(gamma[mu*2]/gamma["m"*2]/"l"[2]),
      expression(mu[0])
    ) %>% setNames(row_names)
    
    col_names <- c("miR-200", "mRNA_PD-L1", "PD-L1_ER")
    col_labels <- c(
      expression("miR-200 (Y"[mu]*")"),
      expression("mRNA PD-L1 (Y"["m"]*")"),
      expression("PD-L1 in ER (L)"[])
    ) %>% setNames(col_names)
    
    par_meta %>% mutate(
      row = factor(row, labels = row_labels),
      col = factor(col, labels = col_labels)
    )
  }),
  tar_target(sens_labels, {
    vars <- c("lmp2", "lmp3", "lmump2", "lmump3", "ltp1", "ltp2", "ltp3", "mu2_0 lmp", "mu2_0 lmump", "mu2_0 ltp")
    c(expression('l'[mP2]),
      expression('l'[mP3]),
      expression('l'[paste(mu,'mP2')]),
      expression('l'[paste(mu,'mP3')]),
      expression('l'[tP1]),
      expression('l'[tP2]),
      expression('l'[tP3]),
      expression(paste(mu['0'], ' ', '(L'['p'['mP']], ')')),
      expression(paste(mu['0'], ' ', '(L'['p'[paste(mu,'mP')]], ')')),
      expression(paste(mu['0'], ' ', '(L'['p'['tP']], ')'))
    ) %>% setNames(vars)
  }),
  tar_target(sens_xmin, 170000 / 1e5),
  tar_target(sens_xmax, 220000 / 1e5)
)

```
```{targets sens_plot}
list(
  tar_target(sens_all, {
    # add model value to all pars (plotting from 2 different data frames I didn't get working)
    sens <- all_nested %>% filter(par != "base")
    base_add <- sens %>% expand(par) %>% add_column(all_nested %>% filter(par == "base") %>% select(-par))
    all <- bind_rows(sens, base_add) %>%
      unnest(cols = data) %>%
      mutate(stable = (interval %% 2) == 0) %>%
      mutate(change = factor(change,
                             levels = c(-0.2, -0.1, 0, 0.1, 0.2),
                             labels = c("-20%", "-10%", "Model value", "+10%", "+20%"),
                             ordered = T))
  }),
  tar_target(sens_pdl_plot, {
    sens_all %>%
      left_join(par_meta, by = "par") %>%
      mutate(par = factor(par, labels = sens_labels)) %>%
      filter(row != "mu[0]") %>%
      ggplot(aes(`# Values[S].InitialValue` / 1e5, `[PM]`, linetype = stable, 
                 color = change, group = interaction(interval, change)))+ 
      geom_line(size = 1) +
      facet_grid(rows = vars(row), cols = vars(col), labeller = label_parsed) + 
      scale_linetype_discrete(guide = "none", limits = c(T,F)) +
      scale_color_discrete_diverging(palette = "Blue-Red") +
      labs(color = NULL,
           x = bquote('SNAIL1 ('*10^5~'molecules)'), 
           y = 'Membrane PD-L1 (molecules)'
      ) +
      theme_cowplot(font_size = text_size) +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_text(size = 1.2*text_size, face = "bold"),
        strip.text.y = element_text(size = 1.2*text_size, face = "bold"),
        legend.position = "bottom",
        legend.justification = "center",
        axis.text.x=element_text(angle = 45, hjust = 1)
      ) + 
      panel_border() + 
      coord_cartesian(xlim = c(sens_xmin,sens_xmax)) +
      scale_y_log10()
  }),
  tar_target(sens_zeb_plot, {
    sens_all %>%
      left_join(par_meta, by = "par") %>%
      mutate(par = factor(par, labels = sens_labels)) %>%
      filter(row != "mu[0]") %>%
      ggplot(aes(`# Values[S].InitialValue` / 1e5, `[Z]`, linetype = stable, color = change, group = interaction(interval, change)))+ 
      geom_line(size = 1) +
      facet_grid(rows = vars(row), cols = vars(col), labeller = label_parsed) + 
      scale_linetype_discrete(guide = "none", limits = c(T,F)) +
      scale_color_discrete_diverging(palette = "Blue-Red") +
      labs(color = NULL,
           x = bquote('SNAIL1 ('*10^5~'molecules)'), 
           y = 'ZEB1 (molecules)'
      ) +
      theme_cowplot(font_size = text_size) +
      #theme_cowplot() + #< for cvd_grid
      theme(
        strip.background = element_blank(),
        strip.text.x = element_text(size = 1.2*text_size, face = "bold"),
        strip.text.y = element_text(size = 1.2*text_size, face = "bold"),
        legend.position = "bottom",
        legend.justification = "center",
        #axis.text.x=element_text(angle = 45, hjust = 1)
      ) + 
      panel_border() + 
      coord_cartesian(xlim = c(sens_xmin,sens_xmax)) +
      scale_y_log10()
  })
)

```

#### Combined

```{targets sens_comb}
list(
  tar_target(sens_comb, {
    plot_grid(
      sens_pdl_plot + theme(axis.title.x=element_blank(), 
                            axis.text.x=element_blank()),
      sens_zeb_plot + theme(legend.position="none",
                            strip.background.x = element_blank(), 
                            strip.text.x = element_blank()),
      ncol = 1)
  }),
  tar_target(sens_comb_file, {
    path <- "output/si_sens_mir_pars.pdf"
    ggsave(path, plot = sens_comb, width = 10, height = 14, device = cairo_pdf)
  })
)
```

### S5. mP * Y dynamics

```{targets si_td_mPY, tar_simple = T}
td_vars <- c("(R5).Flux")
td_labels <- c("mP * Y") %>% setNames(td_vars)

time_dynamics_long %>% 
  filter(name %in% td_vars) %>%
  mutate(experiment = factor(experiment, levels = exp_order)) %>%
  ggplot(aes(time,value * 1e10, color = experiment)) +
  geom_line(aes(linetype = experiment), size = 1.2) +
  facet_grid(rows = vars(factor(name, levels = td_vars)), cols = vars(type), 
             scales = "free_y", switch = "y",
             labeller = labeller(.rows = as_labeller(td_labels))) +
  #facet_wrap(vars(type), labeller = labeller(as_labeller(td_labels))) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  expand_limits(y = 0) + 
  #scale_color_OkabeIto(order = c(8,2,1), labels = exp_labels, name = NULL) +
  #scale_linetype_manual(values = c("dashed", "solid", "solid"), labels = exp_labels, name = NULL) + 
  scale_color_OkabeIto(order = c(8,2,1), labels = exp_labels, name = NULL) +
  scale_linetype_manual(values = c("dashed", "solid", "solid"), labels = exp_labels, name = NULL) + 
  labs(x = "Time (days)", y = bquote('Flux (' * 10^-10 ~ 'molecules/time)')) + 
  theme_cowplot(text_size) + 
  panel_border() +
  #background_grid() + 
  theme(strip.placement = 'outside', 
        strip.background = element_blank(),
        strip.text.x = element_text(size = 1.2*text_size, face = "bold"),
        panel.spacing.x = unit(1, "lines"),
        legend.position = "top",
        legend.justification = "right")
```

#### Combined

```{targets si_td_mPY_file}
  tar_target(si_td_mPY_file, {
    path <- "output/si_temp_dyn_mPY.pdf"
    ggsave(path, plot = si_td_mPY, width = 12, height = 3.5, device = cairo_pdf)
  })
```



### S6. Alt temp dynamics

```{targets temp_dyn_alt_files}
list(
  tar_target(emt_ifn1_alt_file, "data/emt_td/EMT_different_time_dynamics.txt", format = "file"),
  tar_target(met_ifn1_alt_file, "data/emt_td/MET_different_time_dynamics.txt", format = "file")
)
```

```{targets time_dynamics_alt_long, tar_simple = T}
time_dynamics <- tribble(
  ~type, ~experiment, ~file,
  "EMT", "no_int", emt_no_int_file,
  "EMT", "ifn0", emt_ifn0_file,
  "EMT", "ifn1", emt_ifn1_alt_file,
  "MET", "no_int", met_no_int_file,
  "MET", "ifn0", met_ifn0_file,
  "MET", "ifn1", met_ifn1_alt_file,

) %>% 
  rowwise() %>%
  mutate(data = list(read_copasi(file))) %>%
  select(-file) %>% unnest(data)

time_dynamics <- time_dynamics %>%
  mutate(
    across(all_of(c("[mP]", "[PM]")), ~ ifelse(experiment == "no_int", NA, .x)),
    time = `# Time`/24
  ) %>% select(-`# Time`)

time_dynamics %>% 
  pivot_longer(cols = c(where(is.numeric), -time))
```

```{targets figure3_alt, tar_simple = T}
time_dynamics_alt_long %>% 
  filter(name %in% td_vars_to_show) %>%
  mutate(experiment = factor(experiment, levels = exp_order)) %>%
  ggplot(aes(time,value, color = experiment)) +
  geom_line(aes(linetype = experiment), size = 1.2) +
  facet_grid(rows = vars(factor(name, levels = td_vars)), cols = vars(type), 
             scales = "free_y", switch = "y",
             labeller = labeller(.rows = as_labeller(td_labels))) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  expand_limits(y = 0) + 
  scale_color_OkabeIto(order = c(8,2,1), labels = exp_labels, name = NULL) +
  scale_linetype_manual(values = c("dashed", "solid", "solid"), labels = exp_labels, name = NULL) + 
  labs(x = "Time (days)", y = "Regulator (molecules)") + 
  theme_cowplot(text_size) + 
  panel_border() +
  #background_grid() + 
  theme(strip.placement = 'outside', 
        strip.background = element_blank(),
        strip.text.x = element_text(size = 1.2*text_size, face = "bold"),
        panel.spacing.x = unit(1, "lines"),
        legend.position = "top",
        legend.justification = "right")
```

#### Combined

```{targets fig3_alt}
  tar_target(figure3_alt_file, {
    path <- "output/si_temp_dynamics_timescales.pdf"
    ggsave(path, plot = figure3_alt, width = 12, height = 10, device = cairo_pdf)
  })
```

### S7. Jak-Stat steady state

```{targets jakstat-ss}
list(
  tar_target(jakstat_ss_file, "data/jakstat_ss.txt", format = "file"),
  tar_target(jakstat, read_copasi(jakstat_ss_file))
)
```

```{targets jakstat_rename, tar_simple = T}
jakstat %>% 
  select(ifn = `# Values[IFN].InitialValue`, 
         stat = `[x10]`)
```


```{targets jsff_spline, tar_simple = T}
jakstat_for_fit <- jakstat_rename %>% 
  filter(ifn < 0.3)
spline(jakstat_for_fit$ifn, jakstat_for_fit$stat) %>% 
  as_tibble() %>% 
  setNames(c("ifn", "stat"))
```
```{targets g3fit, tar_simple = T}
drc::drm(stat ~ ifn, fct = G.3(), data = jsff_spline)
```
```{targets jakstat_fit, tar_simple = T}
jakstat_rename %>%
  mutate(
    g3 = predict(g3fit, newdata = data.frame(ifn = ifn))
  ) %>%
  pivot_longer(-ifn) %>%
  mutate(name = factor(name, levels = c("stat", "g3"))) %>%
  ggplot(aes(x = ifn, y = value, color = name)) + 
  geom_line(size = 1) +
  theme_cowplot(font_size = text_size) + 
  coord_cartesian(xlim = c(0,0.3)) + 
  scale_color_manual(name = NULL, 
                     labels = c("Steady state", "G.3 fit"), 
                     values = c("black", "red")) +
  labs(x = expression(paste('IFN', gamma, ' (nM)')), 
       y = "STAT1p_2 (nM)") +
  theme_cowplot(font_size = text_size) +
  theme(legend.position = "top",
        legend.justification = "right")
```
#### Combined

```{targets si_jakstat-ss}
  tar_target(jakstat_fit_file, {
    path <- "output/si_jakstat_fit.pdf"
    ggsave(path, plot = jakstat_fit, width = 8, height = 4, device = cairo_pdf)
  })
```


# Pipeline

```{r}
tar_make()
```
